\documentclass[number]{ximera}

%My github password is ximera1
%My GeoGebra password is Ximera1

% set font encoding for PDFLaTeX, XeLaTeX, or LuaTeX
\usepackage{ifxetex,ifluatex}
\newif\ifxetexorluatex
\ifxetex
  \xetexorluatextrue
\else
  \ifluatex
    \xetexorluatextrue
  \else
    \xetexorluatexfalse
  \fi
\fi

\ifxetexorluatex
  \usepackage{fontspec}
\else
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{lmodern}
\fi

\usepackage{hyperref}

\title{General Graphing}
\author{Univ. of Minnesota MathCEP}


% Enable SageTeX to run SageMath code right inside this LaTeX file.
% http://mirrors.ctan.org/macros/latex/contrib/sagetex/sagetexpackage.pdf
% \usepackage{sagetex}

\begin{document}

\begin{abstract}
  In this activity, you will discover the issue involved in using the Law of Sines and Law of Cosines to solve triangles.
\end{abstract}

\maketitle

\begin{enumerate}

\item Translations

\begin{itemize}

\item Group 1: Graph $y = x^2$, $y = x^2 + 4$ and $y = x^2 - 2$

\item Group 2: Graph $y = x$, $y = x+3$ and $y=x-1$

\item Group 3: Graph $y = x^2$, $y = (x-3)^2$ and $y = (x+1)^2$

\item What affect will adding $D$ to a function value have on a graph? What affect will adding $C$ to the input value ($x$) before applying the function have on the graph?

\item Interpret $y = x-1$ and $y = (x-1)$ in two different ways and show that their graphs will be the same.

\item Second task for all groups: 

Graph $y = \displaystyle{\frac{1}{x}}$, $y = \displaystyle{\frac{1}{(x-3)}}$ and $y = \displaystyle{\frac{1}{x} + 2}$

\item If you know what the graph of $y = \sin x$ looks like, can you describe what the graph of $y = (\sin x) + 4$ and $y = \sin (x-\frac{\pi}{4})$ look like?

\end{itemize}

\vskip 0.5 cm

\item Reflections

\begin{itemize}

\item Group 1: Graph $y = e^x$, $y = e^{-x}$, $y=-e^x$ and $y = -e^{-x}$

\item Group 2: Graph $y = \ln x$, $y = \ln (-x)$, $y = - \ln x$ and $y = - \ln (-x)$

\item Group 3: Graph $y = \sqrt x$, $y = \sqrt {-x}$, $y = - \sqrt x$ and $y = - \sqrt {-x}$

\item What affect will placing a negative sign in front of the function value do to the graph? What affect will placing a negative sign on the input value before applying the function have on the graph?

\item Second task for all groups: Graph $y = x^2$, $y = -x^2$, $y = (-x)^2$ and $y = -(-x)^2$

\item Why do the graphs of $y = x^2$ and $y = (-x)^2$ look the same? Give two reason, one by simplifying the second equation algebraically, the second by interpreting the affect of the negative sign on the graph.

\item If you know what the graph of $y = \sin x$ looks like, can you describe what the graph of $y = - \sin x$ and $y = \sin (-x)$ look like?

\end{itemize}

\newpage

\item Magnifications

\begin{itemize}

\item Group 1: Graph $y = \ln x$, $y = 4 \ln x$ and $y = \ln {(4x)}$ (Make special note of where the graph crosses the x-axis.)

\item Group 2: Graph $y = \sqrt x$, $y = 2 \sqrt x$ and $y = \sqrt {2x}$

\item Group 3: Graph $y = \frac{1}{x}$, $y = \frac{2}{x}$ and $y = \frac{1}{2x}$ 

\item What affect will multiplying $A$ to a function value have on a graph? What affect will multiplying $B$ to the input value ($x$) before applying the function have on the graph?

\item If you know what the graph of $y = \sin x$ looks like, can you describe what the graph of $y = A \sin x$ and $y = \sin (Bx)$ look like?

\end{itemize}

\item Summary

Given the graph of a function $y = f(x)$ and the transformed graph $y = \pm A \cdot f(\pm Bx + C) +D$:

\begin{itemize}

\item Which things in the transformation affect the graph horizontally (left and right) and which affect the graph vertically (top and bottom)?

\item How does a multiplier affect the graph? a minus sign? a number added? 

\item If you know the graph of $y = f(x)$, how can you find the graph of $y = -5 f(4x) - 3$? the graph of $y = -5 f(4x + 2) - 3$? 

\end{itemize}

\end{enumerate}


\end{document}
