\documentclass[number]{ximera}

%My github password is ximera1
%My GeoGebra password is Ximera1

\input{../../preamble.tex}

% set font encoding for PDFLaTeX, XeLaTeX, or LuaTeX
%\usepackage{ifxetex,ifluatex}
%\newif\ifxetexorluatex
%\ifxetex
%  \xetexorluatextrue
%\else
%  \ifluatex
%    \xetexorluatextrue
%  \else
%    \xetexorluatexfalse
%  \fi
%\fi
%
%\ifxetexorluatex
%  \usepackage{fontspec}
%\else
%  \usepackage[T1]{fontenc}
%  \usepackage[utf8]{inputenc}
%  \usepackage{lmodern}
%\fi
%
%\usepackage{hyperref}

\title{Solving Systems of Linear Equations}
\author{MathCEP}

% Enable SageTeX to run SageMath code right inside this LaTeX file.
% http://mirrors.ctan.org/macros/latex/contrib/sagetex/sagetexpackage.pdf
% \usepackage{sagetex}




\begin{document}

\begin{abstract}
  In this activity, you will discover formulas for the solutions of $2\times2$ linear systems.
\end{abstract}

\maketitle

Consider the following $2\times2$ linear system of equations. Our goal is to derive an expression for $x$ in terms of the letters $a,b,c,d,p,$ and $q$.
\begin{eqnarray}
ax+by&=&p\\
cx+dy&=&q
\end{eqnarray}

\begin{problem}
Multiply Equation (1) by $d$: $\answer{adx+bdy=pd}$
\end{problem}

\begin{problem}
Multiply Equation (2) by $-b$: $\answer{-bcx-bdy=-bq}$
\end{problem}

\begin{problem}
Add the two new equations from Problems 1 and 2:

$\answer{adx-bcx+ bdy-bdy =pd-bq}$
\begin{hint} You should have two terms whose sum is $0$ in order to eliminate the $y$ variable.  \end{hint}
\end{problem}

\begin{problem}
Solve for $x$ in your equation from Problem 3.

$x=\answer{\frac{pd-bq}{ad-bc}}$

\begin{hint} $adx-bcx=(ad-bc)x$ \end{hint}
\end{problem}

Next, we will solve for $y$ in the same $2\times2$ linear system given by Equations (1) and (2).

\begin{problem}
Multiply Equation (1) by $-c$: $\answer{-acx-bcy=-pc}$
\end{problem}

\begin{problem}
Multiply Equation (2) by $a$: $\answer{acx+ady=qa}$
\end{problem}

\begin{problem}
Add the two new equations from Problems 5 and 6:

$\answer{-acx+acx -bcy+ady=-pc+qa}$
\begin{hint}There should be two terms that cancel out to eliminate the $x$ variable.\end{hint}
\end{problem}

\begin{problem}
Solve for $y$ in your equation from Problem 7:

$y=\answer{\frac{-pc+qa}{-bc+ad}}$
\begin{hint}
$-bcy+ady=(-bc+ad)y$
\end{hint}
\end{problem}

%%%% 3x3

Lastly, we will solve for $x$ in the following $3\times3$ linear system of equations
\begin{eqnarray}
ex+fy+gz&=&r\\
hx+iy+jz&=&s\\
kx+ly+mz&=&t
\end{eqnarray}
by doing the following:

\begin{problem}
Multiply Equation (3) by $-j$: $\answer{-jex-jfy-jgz=-jr}$
\end{problem}

\begin{problem}
Multiply Equation (4) by $g$: $\answer{ghx+giy+gjz=gs}$
\end{problem}

\begin{problem}
Add the two new equations from Problems 9 and 10:

$\answer{-jex+ghx-jfy+giy-jgz+gjz=-jr+gs}$
\begin{hint} There should be two terms that cancel out to eliminate the $z$ variable.\end{hint}
\end{problem}

\begin{problem}
Multiply Equation (5) by $-j$: $\answer{-jkx-jly-jmz=-jt}$
\end{problem}

\begin{problem}
Multiply Equation (4) by $m$: $\answer{mhx+miy+mjz=ms}$
\end{problem}

\begin{problem}
Add the two new equations from Problems 12 and 13:

$\answer{-jkx+mhx-jly+miy-jmz+mjz=-jt+ms}$
\begin{hint} There should be two terms that cancel out to eliminate the $z$ variable.\end{hint}
\end{problem}

\begin{problem}
Your equations in Problems 11 and 14 should be linear equations in the variables $x$ and $y$.
These form a $2\times2$ linear system of equations that you can solve with the equations you came up with in Problems 4 and 8.
\begin{hint}
Your equations for $x$ and $y$ in Problems 4 and 8 were in terms of the variables $a,b,c,d,p,$ and $q$.
In order to apply these formulas the the $2 \times 2$ linear system formed by the equations in Problems 11 and 14, you need to figure out what $a,b,c,d,p,$ and $q$ are equal to in terms of the variables $e,f,g,h,i,j,k,l,m,r,s,$ and $t$.
\end{hint}
\begin{hint}Match your equation from Problem 11 with Equation (1), and match your equation from Problem 14 with Equation (2).\end{hint}
\begin{hint} For example, take $a=(gh-je)$.\end{hint}
\begin{hint}Your answer will be much neater if you leave your terms in factored form, i.e., do not FOIL them out.\end{hint}
\[x=\answer{\frac{(gs-jr)(mi-jl)-(gi-jf)(ms-jt)}{(gh-je)(mi-jl)-(gi-jf)(mh-jk)}}\]
\[y=\answer{\frac{(ms-jt)(gh-je)-(gs-jr)(mh-jk)}{(gh-je)(mi-jl)-(gi-jf)(mh-jk)}}\]
\end{problem}


\end{document}